---
title: "index"
format: html
---

# Data

## NYPD Stop, Question and Frisk Data

This dataset is provided by the [New York Police Department](https://www.nyc.gov/site/nypd/stats/reports-analysis/stopfrisk.page) and contains information about the stop, question, and frisk policy implemented by the NYPD. The data can be found at `/r-data/nypd-stop`

## NYC Crime Data

-   [NYPD Shooting Incident Data (Historic)](): List of every shooting incident that occurred in NYC going back to 2006 through the end of the previous calendar year.
-   [NYPD Complaint Data (Historic)](): List all valid felony, misdemeanor, and violation crimes reported to the New York City Police Department (NYPD) from 2006 to the end of last year
-   [NYPD Arrest Data (Historic)](): List of every arrest in NYC going back to 2006 through the end of the previous calendar year.
-   Misdemeanor Counts by Precinct: List of major misdemeanor crime statistics for each NYPD precinct.
-   Major Felony Counts by Precinct: List of major felony crime statistics for each NYPD precinct.

## NYC GIS Data

-   [NYC Borough Boundaries](): Shapefile data for the five boroughs of New York City.
-   [NYPD Precinct Boundaries](): Shapefile data for the 77 NYPD precincts in New York City.
-   [U.C.S. Census Tracts for New York City](): Shapefile data for the census tracts in New York City.

## NYC Census Data

-   [NYC Census Data by Precinct (2020)](): Census data for New York City broken down by NYPD precincts from the 2020 U.S. Census.

# Setup

`setup.R` contains the code to load and preprocess the datasets used in this analysis -- NYPD stop and crime data, NYC GIS data, and US Census data.

```{r setup, include=FALSE}

# configure data workspace
source("config/load.R")

```

## Missing Value Analysis

```{r missing values analysis}

sqf_hist |>
  st_drop_geometry() |>   
  summarise(across(everything(), ~ sum(is.na(.)))) |>
  pivot_longer(everything(), names_to = "column", values_to = "na_count") |>
  arrange(desc(na_count)) |>
  # filter(na_count == 0) |>
  print()

# NA values for search related fields decrease dramatically when a search is actually conducted
sqf_hist |>
  st_drop_geometry() |> 
  filter(SEARCHED_FLAG == "Y") |>
  # select(matches("SEARCH")) |>
  summarise(across(everything(), ~ sum(is.na(.)))) |>
  pivot_longer(everything(), names_to = "column", values_to = "na_count") |>
  arrange(desc(na_count)) |>
  print()

```

After cleaning the raw dataset and selecting 15 key attributes, the cleaned dataset consists of 15 attributes, including `stop_was_initiated`, `officer_explained_stop_flag`, `other_person_stopped_flag`, `summons_issued_flag`, `officer_in_uniform_flag`, `frisked_flag`, `searched_flag`, `other_contraband_flag`, `weapon_found_flag`, `suspects_actions_decription_flag`, `observed_duration_minutes`, `stop_duration_minutes`, `suspect_sex`, `suspect_reported_age`, and `suspect_race_description.`

The `suspect_sex` and `suspect_race_description` are the protected attributes and are later used to detect and identify mitigation actions.

