---
title: "new_index"
format: html
---

# Data

## NYPD Stop, Question and Frisk Data

This dataset is provided by the [New York Police Department](https://www.nyc.gov/site/nypd/stats/reports-analysis/stopfrisk.page) and contains information about the stop, question, and frisk policy implemented by the NYPD. The data can be found at `/r-data/nypd-stop`

## NYC Crime Data
  - [NYPD Shooting Incident Data (Historic)](): List of every shooting incident that occurred in NYC going back to 2006 through the end of the previous calendar year.
  - [NYPD Complaint Data (Historic)](): List all valid felony, misdemeanor, and violation crimes reported to the New York City Police Department (NYPD) from 2006 to the end of last year
  - [NYPD Arrest Data (Historic)](): List of every arrest in NYC going back to 2006 through the end of the previous calendar year. 
  - Misdemeanor Counts by Precinct: List of major misdemeanor crime statistics for each NYPD precinct.
  - Major Felony Counts by Precinct: List of major felony crime statistics for each NYPD precinct.
  
## NYC GIS Data
  - [NYC Borough Boundaries](): Shapefile data for the five boroughs of New York City.
  - [NYPD Precinct Boundaries](): Shapefile data for the 77 NYPD precincts in New York City.
  - [U.C.S. Census Tracts for New York City](): Shapefile data for the census tracts in New York City.

## NYC Census Data
  - [NYC Census Data by Precinct (2020)](): Census data for New York City broken down by NYPD precincts from the 2020 U.S. Census.

# Setup
`setup.R` contains the code to load and preprocess the datasets used in this analysis -- NYPD stop and crime data, NYC GIS data, and US Census data.


```{r setup, include=FALSE}

# check to see if the setup has already been run, if not, run it
# this includes the NYPD crime and SQF data and R libraries 
if (!exists(".setup_complete") || !.setup_complete) {
      source("setup.R")
    } else {
      message("setup.R has already been run.")
    }

# check to see if the GIS setup has already been run, if not, run it
if (!exists(".setup_complete_gis") || !.setup_complete_gis) {
      source("setup.R")
    } else {
      message("gis.R has already been run.")
    }

# check to see if the Census setup has already been run, if not, run it
# `census.R` depends on `gis.R` being run first
if (!exists(".setup_complete_census") || !.setup_complete_census) {
      source("census.R")
    } else {
      message("census.R has already been run.")
    }

```


# Descriptive Statistics


```{r}

# summary statistics for numeric variables in sqf_hist
sqfSummary <- sqf_hist %>%
  select(where(is.numeric)) %>%
  summary()
print(sqfSummary)


# 

```


# Spatial Analysis

## NYPD Precincts

```{r create map of nypd precincts}

# plot the NYPD map
nypdMAP <- ggplot(nypd_sf) +
  geom_sf(fill = "white", color = "black", linewidth = 0.3) +
  theme_void()

```





