---
title: "OLS Regressions for Arrest Made or Summons Issued"
subtitle: "Controlling for the Detainee's Suspected Crime"
author: "Gabriel Brock"
date: "`r Sys.Date()`"
format:
  pdf:
    documentclass: article
    geometry:
      - margin=1in
---


Table 5 shows the results for arrests, summons, and arrests or summons. 

To test for precinct effects of local crime and social conditions, we estimated the models first as OLS regressions without precinct-fixed effects, and then we re-estimated the models including precinct fixed effects. 

We use the following OLS equation, illustrated here with Sanction as the dependent variable and several controls for suspect characteristics, event (suspected crime) characteristics, local crime rates, and 
social conditions: 

Black, Black Hispanic, and White Hispanic are dummy variables indicating that a person of that race was stopped. 

Model (2) is the same as model (1) except it incorporates additional terms that associate the RS factors with the three race dummy variables. 

The third set of models in Table 5 (in columns 7–9) are two-step arrest models. In this formulation, arrests are a two-stage process: the officer must first choose whether or not the person deserves some sort of sanction. Then, the officer must choose to arrest the person or issue a summons.Accordingly, we estimate arrest as a choice conditional on the decision to sanction. 

We first use a logistic model to estimate the probability of sanction and add the predicted probability of sanction to the estimation of which members of the sanction group are arrested.

# --- Data preparation ----

```{r data prep}

# Using `sqf_ols` from ~/analysis/test-hitrate.qmd, chunk {r stop-level data prep}
# `sqf_ols` is saved to an rds file in that chunk and referenced here for rendering
sqf_ols <- readRDS("~/Projects/thesis/data/derived/sqf_ols.rds")

# load libraries
library(tidyverse)
library(fixest) # for regressions
library(modelsummary) # for printing regression tables

# check for missing values in variables used in models
colSums(is.na(sqf_ols))

# impute zeros for NA values for `sqf_ols$female` 
# (assuming cop would have indicated a female detainee)
sqf_ols$female[is.na(sqf_ols$female)] <- 0

```

```{r partitioning SQF by admin}

# Create data subsets for different mayoral administrations
sqf_bloomberg <- filter(sqf_ols, year >= 2006 & year <= 2013)
sqf_deblasio <- filter(sqf_ols, year >= 2014 & year <= 2021)
sqf_adams <- filter(sqf_ols, year >= 2022)

```


\newpage
# --- Model 1: basic OLS with fixed effects (without interactions) ----

```{r model 1: basic ols}

# Define function to estimate model 1 across different datasets
model1 <- function(data) {
  feols(
    sanction ~ Black + Hisp + age + female + 
      RS_furtive + RS_crimloc + RS_casing + RS_other +
      RS_drug + RS_suspobj + RS_appear + RS_violent | 
      pct + year + pct:year + off_cat_broad,
    data = data,
    cluster = ~ pct
  )
}

# Estimate model 1 for each dataset
model1_all <- model1(sqf_ols)
model1_bloomberg <- model1(sqf_bloomberg)
model1_deblasio <- model1(sqf_deblasio)
model1_adams <- model1(sqf_adams)

```

\newpage
```{r}
etable(
  model1_all, model1_bloomberg, model1_deblasio, model1_adams,
  digits = 3,                 # round coefficients to 3 decimals
  signif.code = c("***"=0.001, "**"=0.01, "*"=0.05, "."=0.1), # custom significance
  order = c("Black","Hisp","age","RS_furtive","RS_crimloc","RS_casing",
            "RS_other","RS_drug","RS_suspobj","RS_appear","RS_violent"), # variable order
  dict = c(
    Black = "Black",
    Hisp = "Hispanic",
    age = "Age",
    RS_furtive = "RS: Furtive",
    RS_crimloc = "RS: Crime Location",
    RS_casing = "RS: Casing",
    RS_other = "RS: Other",
    RS_drug = "RS: Drug",
    RS_suspobj = "RS: Suspicious Object",
    RS_appear = "RS: Appearance",
    RS_violent = "RS: Violent"
  ),
  headers = c(
    "All Stops",
    "Bloomberg Admin.",
    "De Blasio Admin.",
    "Adams Admin."
  ),
  view = TRUE
)
```
\newpage
# --- Model 1b: basic OLS with fixed effects disagreggated by arrests and summons ---

```{r model 1b: basic ols, disggregated by arrest/summons}}

# Define function to estimate model 1 across different datasets

model1b <- function(data) {
  feols(
    c(arrest, summons, sanction) ~ Black + Hisp + age + female + 
      RS_furtive + RS_crimloc + RS_casing + RS_other +
      RS_drug + RS_suspobj + RS_appear + RS_violent | 
      pct + year + pct:year + off_cat_broad,
    data = data,
    cluster = ~ pct
  )
}

# Estimate model 1 for each dataset
model1b_all <- model1b(sqf_ols)
model1b_bloomberg <- model1b(sqf_bloomberg)
model1b_deblasio <- model1b(sqf_deblasio)
model1b_adams <- model1b(sqf_adams)

etable(
    model1b_all,
    model1b_bloomberg,
    model1b_deblasio,
    model1b_adams,
    keep = c("Black", "Hisp"), view = T
)

```


\newpage
# --- Model 2: OLS with (race and suspicion factors) interactions ---

```{r model 2a: race x RS interactions}

#' (A + B + C)*(X + Y + Z) in fixest() expands to main effects plus all 
#' interactions between suspicion factors and race dummies automatically.

model2a <- function(data) { 
  feols(
    sanction ~ 
      RS_furtive + RS_crimloc + RS_casing + RS_other +
          RS_drug + RS_suspobj + RS_appear + RS_violent + 
      Black + Hisp +
      age + female |
      pct + year + pct:year + off_cat_broad, # fixed effects
  data = data,
  cluster = ~pct
 )
} 

# Estimate model 2 for each dataset
model2a_all <- model2a(sqf_ols)
model2a_deblasio <- model2a(sqf_deblasio)
model2a_adams <- model2a(sqf_adams)

etable(
    model2a_all,
    model2a_bloomberg,
    model2a_deblasio,
    model2a_adams,
    order = c("^Black$", "^Hisp$", "^age$", "^female$",
             "^RS_furtive$", "^RS_crimloc$", "^RS_casing$", "^RS_other$",
             "^RS_drug$", "^RS_suspobj$", "^RS_appear$", "^RS_violent$"),
    drop = "x",
    view = TRUE
)

```

```{r model 2b: race x RS interactions}

#' (A + B + C)*(X + Y + Z) in fixest() expands to main effects plus all 
#' interactions between suspicion factors and race dummies automatically.

model2b <- function(data) { 
  feols(
    sanction ~ 
      RS_furtive + RS_crimloc + RS_casing + RS_other +
          RS_drug + RS_suspobj + RS_appear + RS_violent +
      (RS_furtive + RS_crimloc + RS_casing + RS_other +
          RS_drug + RS_suspobj + RS_appear + RS_violent) * Black +
      (RS_furtive + RS_crimloc + RS_casing + RS_other +
          RS_drug + RS_suspobj + RS_appear + RS_violent) * Hisp +
      age + female |
      pct + year + pct:year + off_cat_broad, # fixed effects
  data = data,
  cluster = ~pct
 )
} 

# Estimate model 2 for each dataset
model2b_all <- model2b(sqf_ols)
model2b_bloomberg <- model2b(sqf_bloomberg)
model2b_deblasio <- model2b(sqf_deblasio)
model2b_adams <- model2b(sqf_adams)

etable(
    model2b_all,
    model2b_bloomberg,
    model2b_deblasio,
    model2b_adams,
    order = c("^Black$", "^Hisp$", "^age$", "^female$",
             "^RS_furtive$", "^RS_crimloc$", "^RS_casing$", "^RS_other$",
             "^RS_drug$", "^RS_suspobj$", "^RS_appear$", "^RS_violent$"),
    view = TRUE
)

```

```{r RS/race interactions}

# ---- RS/race interactions ----
mod_sanction_int <- feols(
  sanction ~ Black + Hisp +
    age + female +
    RS_furtive + RS_crimloc + RS_other + RS_suspobj + RS_appear +
    # interactions:
    RS_furtive:Black + RS_furtive:Hisp + 
    RS_crimloc:Black + RS_crimloc:Hisp + 
    RS_other:Black + RS_other:Hisp + 
    RS_suspobj:Black + RS_suspobj:Hisp + 
    RS_appear:Black + RS_appear:Hisp +
    i(off_cat) | pct + year + interaction(pct, year),
  data    = sqf_ols,
  cluster = "pct"
)

summary(mod_sanction_int)

coef_df <- broom::tidy(mod_sanction_int) %>%
  filter(grepl("RS_", term) & grepl(":", term)) %>%
  mutate(
    RS = sub(":.*", "", gsub("RS_", "", term)),
    race = sub(".*:", "", term)
  )

ggplot(coef_df, aes(x = RS, y = estimate, colour = race)) +
  geom_point(position = position_dodge(width = 0.5)) +
  geom_errorbar(aes(ymin = estimate - 1.96*std.error,
                    ymax = estimate + 1.96*std.error),
                width = 0, position = position_dodge(width = 0.5)) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  coord_flip()

```


\newpage
# --- Model 3: Two-stage model for arrest conditional on sanction ----

Modeling officer decision AFTER a stop is made (re: similarly-situated persons):

1. Stage 1: Should this person receive any sanction at all?
2. Stage 2: If a sanction is imposed, should it be arrest rather than a summons?

```{r model 3: arrest cond. sanction}

# Define function to estimate two-stage arrest model
estimate_arrest_model <- function(data) {
  # Step 1: Estimate the probability of sanction (logit)
  sanction_model <- feglm(
    sanction ~ Black + Hisp + age + female + 
      RS_furtive + RS_crimloc + RS_casing + RS_other +
      RS_drug + RS_suspobj + RS_appear + RS_violent |
      pct + year + pct:year + off_cat_broad,
    data = data,
    cluster = ~ pct,
    family = binomial("logit")
  )
  
  # Append predicted probability of sanction to data frame
  data$prob_sanction <- NA_real_
  prob_hat <- predict(sanction_model, type = "response")
  if (length(prob_hat) == nrow(data)) { #' If the nrows of the predicted probs. 
                                        #' equals the nrows in the data, then
    data$prob_sanction <- prob_hat      #' directly assign the predicted probs.
                                        #' to the new column, `prob_sanction`.  
  # Handle missing values in the prediction model
  } else if (!is.null(sanction_model$obs_selection) &&    #' Otherwise, if `obs_selection` exists and has entries,
             length(sanction_model$obs_selection) >= 1) { #' assign the predicted probs. to the selected rows
    idx <- sanction_model$obs_selection[[1]] #' Get the indices of the selected rows
    data$prob_sanction[idx] <- prob_hat      #' Assign the predicted probs. to those rows
  } else {
    stop("Could not align predicted probabilities with input data: model predictions length does not match nrow(data) and obs_selection is unavailable.")
  }
  
  # Sanity check
  sum(is.na(data$prob_sanction)) # NAs should equal dropped obs in Step 1
  summary(data$prob_sanction)
  
    #' `prob_sanction`:
    #' The officer’s predicted propensity to sanction this person, based on 
    #' observable characteristics, suspicion factors, and fixed effects.
  
  # Step 2: Model decision to arrest conditional on sanction
  arrest_model <- feols(
    arrest ~ Black + Hisp + age + female + prob_sanction |
      pct + year + pct:year + off_cat_broad,
    data = subset(data, sanction == 1),
    cluster = ~ pct
  )
  
    #' `arrest_model`:
    #' Given two people who look equally likely to deserve some punishment, 
    #' are officers more likely to arrest one group than another instead of issuing a summons?

  return(arrest_model)
}

# Run model for each dataset
arrest_model_all <- estimate_arrest_model(sqf_ols)
arrest_model_bloomberg <- estimate_arrest_model(sqf_bloomberg)
arrest_model_deblasio <- estimate_arrest_model(sqf_deblasio)
arrest_model_adams <- estimate_arrest_model(sqf_adams)


```

\newpage
```{r}
etable(arrest_model_all, arrest_model_bloomberg, arrest_model_deblasio, arrest_model_adams,
       view = TRUE)
```

Conditioning `arrest_model` on `prob_sanction` serves two primary functions.
    
  First, it controls for selection bias. Due to the increased burden to issue a sanction, officers are more likely to sanction people who appear to be more suspicious or threatening, factors which also affect the likelihood of arrest. Without `prob_sanction`, race coefficients could pick up co-variations in suspicion as Black suspects could be more likely to be arrested because they were more likely to be sanctioned in the first place.
    
  Second, it separates severity choice from sanction choice. Race coefficients are interpreted as racial differences in arrest vs summons, holding constant: observable characteristics, crime type, precinct/year conditions, and the predicted likelihood that a sanction was warranted at all
