---
title: "use-of-force matching"
format: html
---

# Use of Force Analysis

## 1. Data Prep

```{r filter sqf_hist for 2024 data}

# filter sqf_hist for 2024 data only
sqf2024 <- sqf_hist |>
  filter(YEAR2 == 2024)

```

## 2. Matrix Design

```{r data prep for force matrix}

# Step 1: Define force types and their corresponding scores

  # list of types of force
  force_types <- c(
    "PHYSICAL_FORCE_CEW_FLAG",
    "PHYSICAL_FORCE_DRAW_POINT_FIREARM_FLAG",
    "PHYSICAL_FORCE_HANDCUFF_SUSPECT_FLAG",
    "PHYSICAL_FORCE_OC_SPRAY_USED_FLAG",
    "PHYSICAL_FORCE_OTHER_FLAG",
    "PHYSICAL_FORCE_RESTRAINT_USED_FLAG",
    "PHYSICAL_FORCE_VERBAL_INSTRUCTION_FLAG",
    "PHYSICAL_FORCE_WEAPON_IMPACT_FLAG"   
  )

  # vector of corresponding severity and authorization scores for each type of force
  force_scores <- c(3.06,	0.6917, # "PHYSICAL_FORCE_CEW_FLAG"
                    5.46,	0.5312, # "PHYSICAL_FORCE_DRAW_POINT_FIREARM_FLAG"
                    2.11,	0.9346, # "PHYSICAL_FORCE_HANDCUFF_SUSPECT_FLAG"
                    2.64,	0.8198, # "PHYSICAL_FORCE_OC_SPRAY_USED_FLAG"
                    4.40,	0.6748, # "PHYSICAL_FORCE_OTHER_FLAG"
                    2.11,	0.6748, # "PHYSICAL_FORCE_RESTRAINT_USED_FLAG"
                    1.00,	0.9346, # "PHYSICAL_FORCE_VERBAL_INSTRUCTION_FLAG"
                    4.05,	0.5312  # "PHYSICAL_FORCE_WEAPON_IMPACT_FLAG"
                    )

# create force score matrix
force_score_matrix <- matrix(force_scores, 
                             nrow = 8, ncol = 2, 
                             byrow = TRUE) # create a matrix of force scores
rownames(force_score_matrix) <- force_types # assign force types to matrix as row names
colnames(force_score_matrix) <- c("severity", "authorization") # assign column names

```

```{r calculate weighted force scores for stops}

# Step 2: create a matrix of use of force indicators for each unique stop
  # create force indicator matrix
  force_matrix <- sqf2024 |>
    select(STOP_ID, all_of(force_types)) |> # select STOP_ID and all force types
    mutate(across(all_of(force_types), ~ ifelse(. == "Y", 1, 0))) |> # change Y/N to 0/1
    as.data.frame() # turn matrix into data frame
  
  # assign STOP_ID as row names
  rownames(force_matrix) <- force_matrix$STOP_ID # assign stop IDs to matrix as row names
  force_matrix$STOP_ID <- NULL  # remove STOP_ID column

    # convert to matrix
    force_matrix <- as.matrix(force_matrix)
  
    # replace NAs with 0
    force_matrix[is.na(force_matrix)] <- 0

# Step 3: calculate weighted force score for each stop
weighted_score <- force_matrix %*% (force_score_matrix[, "severity"])

  # create data frame of stops with their weighted force scores
  force_weighted_stops <- data.frame(
    STOP_ID = as.numeric(rownames(force_matrix)),
    weighted_score = as.numeric(weighted_score)
  )
  
    # preview weighted force scores for stops
    head(force_weighted_stops)

# Step 4: merge weighted force scores back to sqf2024
sqf2024 <- sqf2024 |>
  left_join(force_weighted_stops, by = "STOP_ID")

```

## 3. Visualize Scores

```{r visualize weighted force scores}

# boxplot of weighted force scores by SUSPECT_RACE_DESCRIPTION
sqf2024 |>
  filter(SUSPECT_RACE_DESCRIPTION %in% BWH) |>
  ggplot(aes(x = SUSPECT_RACE_DESCRIPTION, y = log10(weighted_score + 1))) +
  geom_boxplot() +
  labs(
    x = "Suspect Race",
    y = "Weighted Use-of-Force Score"
  ) +
  theme_minimal()

library(ggplot2)
library(dplyr)

BWH <- c("BLACK", "WHITE", "HISPANIC")  # define races to include

sqf2024 |>
  filter(SUSPECT_RACE_DESCRIPTION %in% BWH) |>
  ggplot(aes(x = SUSPECT_RACE_DESCRIPTION, y = log10(weighted_score))) +
  geom_jitter(width = 0.2, alpha = 0.3, color = "darkblue", size = 0.7) +
  labs(
    x = "Suspect Race",
    y = "Log Weighted Use-of-Force Score)"
  ) +
  theme_minimal()

sqf2024 |>
  filter(SUSPECT_RACE_DESCRIPTION %in% BWH) |>
  ggplot(aes(x = SUSPECT_RACE_DESCRIPTION, y = weighted_score)) +
  geom_jitter(width = 0.2, alpha = 0.3, color = "darkblue", size = 0.7) +
  labs(
    x = "Suspect Race",
    y = "Weighted Use-of-Force Score"
  ) +
  theme_minimal()

```

## 4. Force Statistics

```{r use of force statistics}
# ---
# all stops that involved any use of force
v1 <- sqf2024 |>
  filter(if_any(all_of(force_types), ~ . == "Y")) |>
  count()

  # all stops
  v2 <- count(sqf2024)
  
  # percent of all stops that used force
  v1 / v2 * 100

# ---
# count all force types used in 2024 stops
v3 <- sqf2024 |>
  summarise(across(all_of(force_types), ~ sum(. == "Y", na.rm = TRUE))) |>
  pivot_longer(everything(), names_to = "force_type", values_to = "count")

# count all force types used in 2024 stops
v4 <- sqf2024 |>
  filter(SUSPECT_RACE_DESCRIPTION %in% BWH) |>
  group_by(SUSPECT_RACE_DESCRIPTION) |>
  summarise(across(all_of(force_types), ~ sum(. == "Y", na.rm = TRUE)),
            .groups= "drop")

# ---
# total of each force group by race  
v5 <- sqf2024 |>
  select(SUSPECT_RACE_DESCRIPTION, all_of(force_types)) |>
  group_by(SUSPECT_RACE_DESCRIPTION) |>
  summarize(across(all_of(force_types), 
                   ~ sum(. == "Y", na.rm = TRUE)
                   ), .groups = "drop")

v5 |>
  filter(SUSPECT_RACE_DESCRIPTION %in% BWH) |>
  kbl(format = "html", digits = 3)

# total across force groups by race
v6 <- v5 |>
  select(!c(SUSPECT_RACE_DESCRIPTION, PHYSICAL_FORCE_VERBAL_INSTRUCTION_FLAG)) |>
  rowSums()

v6

# joining back to race group names
v7 <- data.frame(SUSPECT_RACE_DESCRIPTION  = c("AMERICAN INDIAN/ALASKAN NATIVE", "ASIAN / PACIFIC ISLANDER",
                                               "BLACK", "HISPANIC", "MIDDLE EASTERN/SOUTHWEST ASIAN", "WHITE", NA), 
                force_total = v6)

# ---

# stops/force total by race
v8 <- stopsByRace |>
  left_join(v7, by = "SUSPECT_RACE_DESCRIPTION") |>
  rename(stops_total = n)

# perecent of stops that result in force by race
v8 <- v8 |>
  mutate(percent = force_total / stops_total * 100)

# total of all stops that resulted in physical force
sum(v8$force_total) / v2 * 100


```

# Mahalanobis Distance (Stand-Alone)

## 1. Create Dataset for Mahalanobis Distance Calculation

```{r Create Dataset for Mahalanobis Distance Calculation}

# https://www.r-bloggers.com/2021/08/how-to-calculate-mahalanobis-distance-in-r/

# Step 1: Create Dataset
  # select covariates for matching (from Legewie 2016)
  matching_variables_maha <- c("STOP_ID", # stopID for merging later
                          "STOP_FRISK_DATE", "STOP_FRISK_TIME", # datetime variables
                          "STOP_LOCATION_X", "STOP_LOCATION_Y", # location variables
                          "SUSPECT_SEX", "SUSPECT_REPORTED_AGE", "SUSPECT_HEIGHT") # suspect demographics
  
  # select matching variables from sqf2024 for Mahalanobis Distance calculation
  sqf2024_matching_maha <- sqf2024 |>
    select(all_of(matching_variables_maha)) |>
    mutate(STOP_FRISK_DATE = as.numeric(STOP_FRISK_DATE),
           STOP_FRISK_TIME = (as.POSIXct(STOP_FRISK_TIME, format = "%H:%M:%S")),
           STOP_FRISK_TIME = as.numeric(STOP_FRISK_TIME),
           SUSPECT_SEX = if_else(SUSPECT_SEX == "MALE", 1, 0))

  str(sqf2024_matching_maha)
  
```

## 2. Calculate Mahalanobis Distance

Syntax from `mahalanobis()` function mentioned as follows, `mahalanobis(x, center, cov)` where: - `x`: indicate matrix of data - `center`: indicate the mean vector of the distribution - `cov`: indicate the [covariance matrix of the distribution](https://datascienceplus.com/understanding-the-covariance-matrix/)

```{r For each observation calculate the Mahalanobis distance}
# Step 2: For each observation calculate the Mahalanobis distance
  
sqf2024_matching_maha$mahalanobis <- mahalanobis(sqf2024_matching_maha, 
                                            colMeans(sqf2024_matching_maha, na.rm = TRUE), 
                                            cov(sqf2024_matching_maha, use = "complete.obs"))

```

## 3. Calculate p-value for each observation

Based on the step 2 result, some of the distances are much higher than others. We want to identify any of the distances that are statistically significant (so we calculate $p$-value).

The $p$-value for each distance is calculated as the Chi-Square statistic of the Mahalanobis distance with $k-1$ degrees of freedom, where $k$ is the number of variables.

In general, a p-value that is less than $0.001$ is considered to be an outlier. In this case, all the $p$-value are greater than $0.001$.

```{r}

# Step 3: Calculate the p-value
# We want to identify any of the distances that are statistically significant (so we calculate p-values).

sqf2024_matching_maha$pvalue <- pchisq(sqf2024_matching_maha$mahalanobis, df=3, lower.tail=FALSE)

# View the results
slice_min(sqf2024_matching_maha, order_by = pvalue, n = 100)

sqf2024_matching_maha |>
  filter(pvalue < 0.001) |>
  nrow()


```

# Matching Procedure

https://cran.r-project.org/web/packages/MatchIt/vignettes/MatchIt.html#ref-ho2007

The main analysis relies on determining the actual use of force statistics for a given precinct $BeforeEvent$ and $AfterEvent$ a boundary shooting incident, then compare the actual trend to a constructed counterfactual $AfterEvent$ trend. To do this, we need to create a matched dataset of stops that are similar in nature to those that involved a boundary shooting.

Matching is used to estimating the causal effect of a binary exposure (i.e., a shooting) on an outcome (i.e., use-of-force proportions) while controlling for measured pre-treatment variables (confounding variables). The goal of matching is to produce covariate balance, that is, for the distributions of covariates in the two groups to be approximately equal to each other, as they would be in a successful randomized experiment.

A matching analysis involves four primary steps: 1) *planning*, 2) *matching*, 3) *assessing the quality of matches*, and 4) *estimating the treatment effect* and its uncertainty.

## Planning

Planning involves

-   selecting the type of effect to be estimated,
-   selecting the target population to which the treatment effect is to generalize, and
-   selecting the covariates for which balance is required for an unbiased estimate of the treatment effect.

The matching procedure for the boundary shooting analysis is designed to estimate the average treatment effect on the treated (ATT), that is, the average effect of a shooting (`treat`) on use-of-force (`weighted_score`) among stops that occurred $AfterEvent$ a shooting. The target population is all stops that occurred in a precinct $AfterEvent$ a shooting.

This matching procedure relies on three decisions, determining:

1.  The covariates to be used for matching.

    -   `treat` *binary* (new variable indicating whether the stop is in the $AfterEvent$ period)
    -   `STOP_LOCATION_PRECINCT` *num* (used for exact matching)
    -   `SUSPECT_RACE_DESCRIPTION` *chr* (used for exact matching)
    -   `STOP_LOCATION` *num* (mahalanobis distance based on X and Y coordinates),
    -   `STOP_FRISK_DATE` (clock circle coordinates),
    -   `SUSPECT_SEX` *binary*
    -   `SUSPECT_REPORTED_AGE` *num*
    -   `SUSPECT_HEIGHT` *num*

2.  The ratio of $BeforeEvent$ stops to match to any given $AfterEvent$ stop.

    -   $1:5$, $1:10$, $1:15$, and $1:25$

3.  The size of the time interval for the $BeforeEvent$.

    -   30 days, 90 days, 180 days, and 365 days

### 1. Data Setup

```{r include = FALSE}

boundary_points |>
  filter(YEAR == 2024,
         PERP_RACE == "BLACK",
         STATISTICAL_MURDER_FLAG == T)

```

```{r create dataset of stops}

# Step 1: create dataset of stops
  # select covariates for matching (from Legewie 2016)
  matching_variables <- c("STOP_ID", # stopID for merging later
                          "STOP_FRISK_DATE", "STOP_FRISK_TIME", # datetime variables
                          "STOP_LOCATION_X", "STOP_LOCATION_Y", # location variables
                          # "STOP_LOCATION_PRECINCT", # -- save for exact matching
                          # "SUSPECT_RACE_DESCRIPTION", # -- save for exact matching
                          "SUSPECT_SEX", "SUSPECT_REPORTED_AGE", "SUSPECT_HEIGHT") # suspect demographics

  # select matching variables from sqf2024 for Mahalanobis Distance calculation
    sqf2024_matching <- sqf2024 |>
      select(all_of(matching_variables)) |>
      mutate(STOP_FRISK_DATE = as.numeric(STOP_FRISK_DATE),
             STOP_FRISK_TIME = (as.POSIXct(STOP_FRISK_TIME, format = "%H:%M:%S")),
             STOP_FRISK_TIME = as.numeric(STOP_FRISK_TIME),
             SUSPECT_SEX = if_else(SUSPECT_SEX == "MALE", 1, 0))
  
    str(sqf2024_matching)

```

```{r create data set of shootings}

# Step 2: create dataset of shootings
  # create list of boundary shooting event dates
  boundary_points |> # referencing `boundary_points` from earlier analysis
    select(INCIDENT_KEY, OCCUR_DATE, PRECINCT) |>
    arrange(desc(OCCUR_DATE)) |>
    filter(OCCUR_DATE >= "2018-01-01") # filter for shootings since we have detailed SQF data


```

### 2. Define time windows for control pool

Create multiple datasets using different pre-event intervals (e.g., 1 month to 1 year).

```{r make treated variable for stops}

# after validating analysis, I need to modify this to loop through each shooting date
# [for testing] sample boundary shooting date
  shooting_date <- as.Date("2024-07-30")

# Step 3: create `treated` variable for stops, relative to shooting date
  # 3.1 create stop variable indicating treatment group relative to shooting
  make_treated <- function(data, shooting_date) {
    
    data |>
      mutate(
        treated = if_else(STOP_FRISK_DATE %within% 
                            interval(shooting_date, shooting_date %m+% weeks(2)), 1, 0),
        event_date = shooting_date
      )
  }

    # test `treated` function
    make_treated(sqf2024, shooting_date) |>
      select(STOP_FRISK_DATE, treated, event_date) |>
      slice_sample(n = 50) # check random sample for 1s and 0s
    
```

```{r make treatment time window for stops}

  # 3.2 create multiple datasets using different pre-event intervals
  # Either 1 month, 3 months, 6 months, or 12 months back for $BeforeEvent$ group
  # Always all stops withing 2 weeks for $AfterEvent$ group
  make_window <- function(data, months_back) {
    window <- interval(ymd(shooting_date), ymd(shooting_date) %m+% weeks(2))
    data |>
      filter(
        (treated == 1 & STOP_FRISK_DATE %within% window |
        (treated == 0 & STOP_FRISK_DATE >= event_date %m-% months(months_back))
      ))
  }

    # test window function
    make_window(make_treated(sqf2024, shooting_date), 1) |>
      select(STOP_FRISK_DATE, STOP_FRISK_TIME, treated, event_date) |>
      slice_sample(n = 50) |> # check random sample for dates
      arrange(desc(STOP_FRISK_DATE))

```

### 3. Transform time of day into circular coordinates

[https://towardsdatascience.com/cyclical-encoding-an-alternative-to-one-hot-encoding-for-time-series-features-4db46248ebba/]()

Time series features are cyclical by nature. When the clock strikes 24:00 (12AM), a new day begins and the next hour is 1:00 (1AM). Although the numbers 1 and 24 are actually the furthest apart number wise, 1 is as close to 24 as 23 is because they are in a cycle. Cyclical encoding transforms these features into a format that represents how some values are closer together (eg 12AM and 1AM) and others are further apart (eg 12AM and 12PM).

By mapping a timestamp on the unit circle via sine/cosine coordinates, we can represent time as it would exist on a clock circle. Think of the right side of the circle as the starting point (It is denoted by a 0 on the chart below) or 00:00 (12AM) on a real 24 hour timescale, which we will divide into 4 6-hour landmarks in order to be able to map hours onto the circle.

As you move counterclockwise up the unit circle, it increases to $\frac{\pi}{2}$ (or $90\deg$ 90 degrees) which would be the equivalent of 6:00AM, $\pi$ ($180 \deg$) or 12:00PM, $\frac{3\pi}{2}$ or 6:00PM, and finally back to $0$ at 12AM. Each time point in between any of those landmarks has its own unique coordinates. In this way we can represent the 24-hour daily cycle using sine and cosine.

```{r transform time into circular coordinates for matching}

# Step 4: transform time into continuous variable (circular coordinates) for matching
  # 4.1 indicate day/night (in case I need to consider this re: Veil of Darkness)
  day_night <- function(data) {
    data |>
      mutate(
       STOP_FRISK_TIME = as.POSIXct(STOP_FRISK_TIME, format = "%H:%M:%S"),
       day_night = if_else(pm(STOP_FRISK_TIME), 1, 0 # check if time is in the PM, if yes then 1 (night)
         )
       )
  }

    # test day/night function
    day_night(sqf2024) |>
      select(STOP_FRISK_TIME, day_night) |>
      slice_sample(n = 50) # check random sample for 0s and 1s

  # 4.2 convert time to clock circle coordinates
    trans_to_clock_time <- function(data) {
      # find number for seconds in a day 
      day = 24*60*60
      
      # transform to clock time
      data |>
        # convert STOP_FRISK_TIME to numerical variable (unix time)
        mutate(STOP_FRISK_TIME = as.POSIXct(STOP_FRISK_TIME, format = "%H:%M:%S"),
          secs_midnight = hour(STOP_FRISK_TIME) * 3600 +
                      minute(STOP_FRISK_TIME) * 60 +
                      second(STOP_FRISK_TIME),
               time_x = sin(secs_midnight * 2 * pi / day),
               time_y = cos(secs_midnight * 2 * pi / day)
    )
}
    
    # test clock time function
    trans_to_clock_time(sqf2024) |>
      select(STOP_FRISK_TIME, time_x, time_y) |>
      slice_sample(n = 50) # check random sample for time_x and time_y
  

```

### 4. Check Initial Imbalance

After planning and prior to matching, it can be a good idea to view the initial imbalance in one’s data that matching is attempting to eliminate.

```{r assemble stops df for match testing}

# select variables for matching (from Legewie 2016)
matching_variables <- c("STOP_ID", # ID for unique identification
                        "STOP_FRISK_DATE", "treated", "time_x", 'time_y', "day_night", # datetime variables
                        "STOP_LOCATION_PRECINCT", "STOP_LOCATION_X", "STOP_LOCATION_Y", # location variables
                        "SUSPECT_RACE_DESCRIPTION", "SUSPECT_SEX", "SUSPECT_REPORTED_AGE", "SUSPECT_HEIGHT" # suspect demographics
                        # "weighted_score" # outcome variable
                        ) 
```


```{r}
sqf2024_matching <- sqf2024 |>
  make_treated(shooting_date) |>
  make_window(12) |>
  day_night() |>
  trans_to_clock_time() |>
  select(all_of(matching_variables)) |>
  mutate(SUSPECT_SEX = if_else(SUSPECT_SEX == "MALE", 1, 0)) |>
  filter(SUSPECT_RACE_DESCRIPTION %in% BWH) |>
  drop_na(all_of(matching_variables))
  
head(sqf2024_matching, n = 25)

```

```{r check initial balance}

library(MatchIt)

# Typically, the method argument specifies the method of matching to be performed; here, we set it to NULL so we can assess balance prior to matching.

m.out0 <- matchit(treated ~ STOP_LOCATION_X + STOP_LOCATION_Y +
                          time_x + time_y +
                          SUSPECT_SEX + SUSPECT_REPORTED_AGE + SUSPECT_HEIGHT,
                  data = sqf2024_matching,
                  method = NULL,
                  distance = "mahalanobis",
                  exact = ~ STOP_LOCATION_PRECINCT)

m.out0

summary(m.out0)

```

## Matching

```{r perform stop matching}

m.out1 <- matchit(treated ~ STOP_LOCATION_X + STOP_LOCATION_Y +
                    time_x + time_y +
                    SUSPECT_SEX + SUSPECT_REPORTED_AGE + SUSPECT_HEIGHT,
                  data = sqf2024_matching,
                  method = "nearest", 
                  distance = "glm",        # This creates the propensity score for plotting
                  mahvars = ~ STOP_LOCATION_X + STOP_LOCATION_Y + 
                              time_x + time_y + SUSPECT_REPORTED_AGE, # This does the matching
                  replace = FALSE,
                  exact = ~ STOP_LOCATION_PRECINCT + SUSPECT_RACE_DESCRIPTION)

m.out1

summary(m.out1)

```

The matching outputs are contained in the `m.out1` object. The key components of the m.out1 object are 
- `weights` (the computed matching weights), 
- `subclass` (matching pair membership), 
- `distance` (the estimated propensity score), and 
- `match.matrix` (which control units are matched to each treated unit). 

How these can be used for estimating the effect of the treatment after matching is detailed in `vignette("estimating-effects")`.

## Assessing the Quality of Matches

In addition to covariate balance, the quality of the match is determined by how many units remain after matching. 

Matching often involves discarding units that are not paired with other units. If, after matching, the remaining sample size is small, the resulting effect estimate may be imprecise. Prospective power analyses can be used to determine how small a sample can be before necessary precision is sacrificed.

To assess the quality of the resulting matches numerically, we can use the summary() function on m.out1 as before. Here we set un = FALSE to suppress display of the balance before matching for brevity and because we already saw it. (Leaving it as TRUE, its default, would display balance both before and after matching.

```{r}

# Checking balance after mahalanobis matching
summary(m.out1, un = FALSE)

```

```{r}

m.out2 <- matchit(treated ~ STOP_LOCATION_X + STOP_LOCATION_Y +
                    time_x + time_y +
                    SUSPECT_SEX + SUSPECT_REPORTED_AGE + SUSPECT_HEIGHT,
                  data = sqf2024_matching,
                  method = "optimal", 
                  distance = "glm",        # This creates the propensity score for plotting
                  mahvars = ~ STOP_LOCATION_X + STOP_LOCATION_Y + 
                              time_x + time_y + SUSPECT_REPORTED_AGE + SUSPECT_SEX, # This does the matching
                  replace = FALSE,
                  exact = ~ STOP_LOCATION_PRECINCT + SUSPECT_RACE_DESCRIPTION)

```



```{r}

plot(m.out1, type = "jitter")
plot(m.out2, type = "jitter")

# create love plot to visualize covariate balance
plot(summary(m.out1))
plot(summary(m.out2))

```

```{r extract the matched stop dataset}

# extract the matched dataset
# dataset only contains the matched units and adds columns for distance, weights, and subclass
m.data <- match_data(m.out2)

head(m.data)

```

```{r}

# prepare the matched data for hit rates
m.data <- m.data %>%
  left_join(sqf_hist %>% select(STOP_ID, FIREARM_FLAG), by = "STOP_ID") %>%
  mutate(hit = if_else(FIREARM_FLAG == "Y", 1, 0))

```


We can then model the outcome in this dataset using the standard regression functions in R, like `lm()` or `glm()`, being sure to include the matching weights (stored in the `weights` variable of the `match_data()` output) in the estimation.

Finally, we use `marginaleffects::avg_comparisons()` to perform $g$-computation to estimate the ATT. We recommend using cluster-robust standard errors for most analyses, with pair membership as the clustering variable; `avg_comparisons()` makes this straightforward.


```{r model the outcome, use-of-force weighted score}}

library("marginaleffects")

# model the outcome, use-of-force weighted score
fit <- lm(
  weighted_score ~ treated,
  data = m.data,
  weights = weights
)

summary(fit)

avg_comparisons(
  fit,
  variables = "treated",
  vcov = ~ subclass,
  newdata = subset(m.data, treated == 1)
)

```

