---
title: "test-disparity"
format: pdf
---

1. Data structure and preprocessing  
2. Computing the spatial “local hit rate” feature  
3. Fitting the big logistic model with interactions  
4. Generating ex ante probabilities and doing the key analyses  
5. “Similarly situated white” counterfactual  
6. Ranking stops and efficiency curves  
7. Building the simplified heuristic model with sign constraints  
8. Applying the heuristic in practice

## 1. Data structure and basic preprocessing

Assume a data frame `stops` with (at minimum):

```r
str(stops)
# id                : unique stop id
# date              : Date
# time              : hh:mm or POSIXct
# year              : numeric
# precinct          : factor or integer 1–123
# lon, lat          : numeric, WGS84
# location_type     : factor (c("housing", "transit", "neither"))
# inside            : factor (c("inside","outside"))
# race              : factor (c("white","black","hispanic","asian","other"))
# sex               : factor (c("male","female"))
# build             : factor (c("heavy","medium","muscular","thin"))
# age, height, weight, obs_time: numeric
# radio_run         : 0/1
# in_uniform        : 0/1
# reason_*          : 0/1 indicators for each UF-250 “stop circumstance”
# suspected_crime   : factor (e.g. "CPW", "drugs", ...)
# weapon_found      : 0/1 (any weapon)
# gun_found         : 0/1
```

```{r Setup and cleaning}
library(dplyr)
library(lubridate)
library(glmnet)
library(ggplot2)
```

```{r variable creation}

sqf_all_clean <- sqf_all %>%
  mutate(
    # outcome
    weapon_found = if_else(WEAPON_FOUND_FLAG == "Y", 1L, 0L, missing = 0L),

    # suspected crime
    suspected_crime = SUSPECTED_CRIME_DESCRIPTION,

    # race (simplify labels as needed)
    race = case_when(
      SUSPECT_RACE_DESCRIPTION %in% c("WHITE", "WHITE HISPANIC") ~ "WHITE",
      SUSPECT_RACE_DESCRIPTION %in% c("BLACK", "BLACK HISPANIC") ~ "BLACK",
      SUSPECT_RACE_DESCRIPTION %in% c("ASIAN / PACIFIC ISLANDER") ~ "ASIAN",
      TRUE ~ "OTHER"
    ),
    race = factor(race),

    sex   = factor(SUSPECT_SEX),
    age   = SUSPECT_REPORTED_AGE,

    precinct = factor(STOP_LOCATION_PRECINCT),

    # date/time fields
    date = STOP_FRISK_DATE,
    year = YEAR2,
    month = factor(MONTH2, levels = 1:12,
                   labels = month.abb[1:12]),
    # if DAY2 is day-of-month, derive weekday from date
    dow   = factor(lubridate::wday(date, label = TRUE)),

    # 4-hour time-of-day bins
    time_parsed = hm(STOP_FRISK_TIME),         
    hour = if_else(!is.na(time_parsed),
                   hour(time_parsed), NA_integer_),
    tod4 = cut(hour,
               breaks = c(0,4,8,12,16,20,24),
               include.lowest = TRUE,
               labels = c("0-4","4-8","8-12","12-16","16-20","20-24"))
  )

```

```{r Create 0/1 predictors from flags}

sqf_all_clean <- sqf_all_clean %>%
  mutate(
    # stop reasons:
    reason_actions_desc            = if_else(SUSPECTS_ACTIONS_DECRIPTION_FLAG == "Y", 1L, 0L, missing = 0L),
    reason_proximity_scene         = if_else(SUSPECTS_ACTIONS_PROXIMITY_TO_SCENE_FLAG == "Y", 1L, 0L, missing = 0L),
    reason_casing                  = if_else(SUSPECTS_ACTIONS_CASING_FLAG == "Y", 1L, 0L, missing = 0L),
    reason_lookout                 = if_else(SUSPECTS_ACTIONS_LOOKOUT_FLAG == "Y", 1L, 0L, missing = 0L),
    reason_drug_transaction        = if_else(SUSPECTS_ACTIONS_DRUG_TRANSACTIONS_FLAG == "Y", 1L, 0L, missing = 0L),
    reason_concealed_weapon        = if_else(SUSPECTS_ACTIONS_CONCEALED_POSSESSION_WEAPON_FLAG == "Y", 1L, 0L, missing = 0L),
    reason_known_to_carry_weapon   = if_else(BACKROUND_CIRCUMSTANCES_SUSPECT_KNOWN_TO_CARRY_WEAPON_FLAG == "Y", 1L, 0L, missing = 0L),
    reason_violent_crime           = if_else(BACKROUND_CIRCUMSTANCES_VIOLENT_CRIME_FLAG == "Y", 1L, 0L, missing = 0L)
  )

```

# 2. Focus on CPW and compute previous year precinct hit rate

```{r Subset to CPW}

sqf_cpw <- sqf_all_clean %>%
  filter(suspected_crime == "CPW") %>%
  filter(!is.na(year), !is.na(precinct))

```

```{r Year-lagged precinct hit rate}

#' For each year-precinct, compute the average weapon hit rate
#' assign as local_hit_prev for the next year's stops in that precinct

# per-year, per-precinct hit rate
precinct_year_hit <- sqf_cpw %>%
  group_by(year, precinct) %>%
  summarize(hit_rate = mean(weapon_found), .groups = "drop")

# lag by 1 year
precinct_year_hit_prev <- precinct_year_hit %>%
  mutate(year_next = year + 1) %>%
  select(year_next, precinct, hit_rate_prev = hit_rate)

# join back onto CPW data
sqf_cpw <- sqf_cpw %>%
  left_join(precinct_year_hit_prev,
            by = c("year" = "year_next", "precinct")) %>%
  mutate(
    local_hit_prev = if_else(is.na(hit_rate_prev), 0, hit_rate_prev)
  ) %>%
  select(-hit_rate_prev)
```

# 3. Train/test split

```{r Train/test split}

sort(unique(sqf_cpw$year))
# see what you actually have; then pick ranges, e.g.:

train <- sqf_cpw %>%
  filter(year %in% c(2010, 2011))  

test  <- sqf_cpw %>%
  filter(year %in% c(2012, 2012)) 

```

# 4. Build a big logistic model with interactions

```{r choose predictors}

stop_reasons <- c(
  "reason_actions_desc",
  "reason_proximity_scene",
  "reason_casing",
  "reason_lookout",
  "reason_drug_transaction",
  "reason_concealed_weapon",
  "reason_known_to_carry_weapon",
  "reason_violent_crime"
)

vars_main <- c(
  "race", "sex",
  "age",
  "precinct",
  "month", "dow", "tod4",
  "local_hit_prev",
  stop_reasons
)

```

```{r build model matrix with pairwise interactions}

# Check which variables actually exist in train data
existing_vars <- intersect(vars_main, names(train))
cat("Variables found in train data:", length(existing_vars), "out of", length(vars_main), "\n")
cat("Missing variables:", setdiff(vars_main, names(train)), "\n")

# Check train dataframe itself
cat("Train dataframe dimensions:", dim(train), "\n")
cat("Train dataframe row count:", nrow(train), "\n")
if (nrow(train) > 0) {
  cat("First few column names:", head(names(train), 10), "\n")
} else {
  cat("Train dataframe is empty!\n")
  stop("Train dataframe has no rows")
}

# Use only existing variables and remove those with all NAs
vars_main_actual <- existing_vars

# Remove variables with all NAs
vars_main_actual <- vars_main_actual[!sapply(vars_main_actual, function(var) {
  all(is.na(train[[var]]))
})]

cat("Variables after removing all-NA columns:", vars_main_actual, "\n")

train_mod <- train %>%
  select(all_of(vars_main_actual))

# Remove rows with any remaining NAs to avoid model.matrix issues
train_mod <- train_mod %>%
  filter(complete.cases(.))

cat("Train_mod dimensions after removing NAs:", dim(train_mod), "\n")

# formula with all 2-way interactions, minus intercept
fmla <- as.formula(
  paste("~ (", paste(vars_main_actual, collapse = " + "), ")^2 - 1")
)

# Check for issues in train_mod
cat("Checking for missing values in each column:\n")
for (col in names(train_mod)) {
  na_count <- sum(is.na(train_mod[[col]]))
  if (na_count > 0) {
    cat(col, ":", na_count, "NAs\n")
  }
}

cat("\nChecking factor levels:\n")
for (col in names(train_mod)) {
  if (is.factor(train_mod[[col]])) {
    cat(col, ":", nlevels(train_mod[[col]]), "levels\n")
    if (nlevels(train_mod[[col]]) == 0) {
      cat("  -> WARNING: Factor with 0 levels!\n")
    }
  }
}

# Try a simpler formula first
simple_fmla <- as.formula(paste("~", paste(vars_main_actual, collapse = " + ")))
cat("\nTrying simple formula without interactions...\n")
X_train_simple <- model.matrix(simple_fmla, data = train_mod)
cat("Simple model matrix dimensions:", dim(X_train_simple), "\n")

# If simple works, try interactions (but limit to avoid memory issues)
if (nrow(X_train_simple) > 0) {
  cat("\nSimple formula worked, but interactions are too large\n")
  cat("Using simple model without interactions to avoid memory issues\n")
  X_train <- X_train_simple
} else {
  cat("\nSimple formula failed\n")
  X_train <- X_train_simple
}

# Update response variable to match filtered data
train_indices <- complete.cases(train_mod)
y_train <- train$weapon_found[train_indices]
cat("Response vector length after filtering:", length(y_train), "\n")
cat("X_train rows:", nrow(X_train), "y_train length:", length(y_train), "\n")

```

```{r fit penalized logistic regression}

fit_glmnet <- glmnet(
  x = X_train,
  y = y_train,
  family   = "binomial",
  alpha    = 0,    # ridge (like strong L2 regularization)
  standardize = TRUE
)

cv_fit <- cv.glmnet(
  x = X_train,
  y = y_train,
  family = "binomial",
  alpha  = 0,
  nfolds = 5
)

lambda_star <- cv_fit$lambda.min

```

# 5. Ex ante probabilities on test set

```{r}

X_test <- model.matrix(fmla, data = test)

test$pred_prob <- as.numeric(
  predict(cv_fit, newx = X_test, s = lambda_star, type = "response")
)

# test$pred_prob is your ex‑ante hit rate estimate for each text CPW stop

```

```{r summaries}

# distribution overall
mean(test$weapon_found)            # overall empirical hit rate
mean(test$pred_prob < 0.01)        # fraction with <1% chance
mean(test$pred_prob < 0.005)       # fraction with <0.5% chance

# by race
test %>%
  group_by(race) %>%
  summarize(
    emp_hit   = mean(weapon_found),
    pred_hit  = mean(pred_prob),
    frac_lt1  = mean(pred_prob < 0.01),
    n = n(),
    .groups = "drop"
  )

```

```{r efficiency curve, prop. stops vs prop weapons}

test_ranked <- test %>%
  arrange(desc(pred_prob)) %>%
  mutate(
    rank         = row_number(),
    frac_stops   = rank / n(),
    cum_weapons  = cumsum(weapon_found),
    frac_weapons = cum_weapons / sum(weapon_found)
  )

ggplot(test_ranked, aes(x = frac_stops, y = frac_weapons)) +
  geom_line() +
  scale_x_log10() +
  labs(x = "Fraction of stops (log scale)",
       y = "Fraction of weapons recovered")

```

# 6. Similarly situated white counterfactuals

```{r}

black_test <- test %>% filter(race == "BLACK")

as_if_white <- black_test
as_if_white$race <- factor("WHITE", levels = levels(test$race))

X_black       <- model.matrix(fmla, data = black_test)
X_as_if_white <- model.matrix(fmla, data = as_if_white)

p_black       <- as.numeric(predict(cv_fit, newx = X_black,       s = lambda_star, type = "response"))
p_as_if_white <- as.numeric(predict(cv_fit, newx = X_as_if_white, s = lambda_star, type = "response"))

data.frame(
  group = c("black_empirical", "as_if_white_pred"),
  hit   = c(mean(black_test$weapon_found), mean(p_as_if_white))
)

```

